Forcing for Isabelle

### 2018-12-28 LICS/intro_m.tex
cosas menores

After G\"odel's Incompleteness Theorems, we cannot expect to have a
formal proof of the consistency of Set Theory in $ZF$. Besides its own
consistency, there are other results which are undecided by $ZF$: the
undecidability of Continuum Hypothesis leads to the development of
techniques for independence proofs. First G\"odel introduced the
theory of \emph{inner models}, which gives rise to his model $L$ of
the \emph{Axiom of Constructibility} \cite{godel-L} and proved the
relative consistency of the Axiom of Choice and the Generalized
Continuum Hypothesis with $\ZF$. Thirty years later Paul
J. Cohen~\cite{Cohen-CH-PNAS} devised the technique of \emph{forcing},
which is the only known way of \emph{extending} models of $\ZF$; in
particular, it can be used to prove the relative consistency of the
negation of the Continuum Hypothesis.

>>>>

After G\"odel's Incompleteness Theorems, we cannot expect to have a
formal proof of the consistency of Set Theory in $\ZF$. Besides its own
consistency, there are other results which are undecided by $\ZF$: the
undecidability of Continuum Hypothesis leads to the development of
techniques for independence proofs. First G\"odel introduced the
theory of \emph{inner models}, which gives rise to his model $L$ of
the \emph{Axiom of Constructibility} \cite{godel-L} and proved the
relative consistency of the Axiom of Choice and the Generalized
Continuum Hypothesis with $\ZF$. Thirty years later Paul
J. Cohen~\cite{Cohen-CH-PNAS} devised the technique of \emph{forcing},
which is the only known way of \emph{extending} models of $\ZF$; in
particular, it can be used to prove the relative consistency of the
negation of the Continuum Hypothesis.
